### 2012-06-30 4.tex
Feverish writing on chapters 2-4

%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter will appear in the proceedings of the 30th West Coast Conference on Formal Linguistics under the same name.}}
\label{clusters}

@@

Reconsider the null hypothesis presented in the preceding chapter:

@@

One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
This technique has been applied to consonants in Semitic, especially Arabic \citep{McCarthy1988,McCarthy1994,Pierrehumbert1993,Frisch1996,Frisch2004,Coetzee2008a} but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English \citep{Berkley1994b,Berkley1994a,Pierrehumbert1994,Dmitrieva2008a,Dmitrieva2008b,Coetzee2008b}, Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.

@@

\section{Domain}

@@

@@

@@

\citet{Aronoff1976}

\citet{Harley2009}

\citet{Taft1975} ?
\citet{Taft1981} ?
\citet{Taft2004a} ?

\citet{Halle1962}

% 2.1.4: k

Under the parses and phonologization described immediately below, the 6,876 simplex CELEX entries contain 21 medial coda types and 40 unique medial onsets. Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, a saturation rate of 18.8\%.

\subsection{Syllabification}

For this study, it is necessary to separate medial consonant clusters into codas and onsets. While CELEX provides syllabified transcriptions, no description of the syllabification procedure is given in the documentation, and inspection of these syllabifications suggests the procedure used is not fully systematic. For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][294]{AOE2}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast, and many other such contrasts in the CELEX data, are dubious simply because there is no clear evidence that contrastive syllabification occurs in any language. Apparent counterexamples appear to be effects of vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which explain the \emph{chemistry}/\emph{ministry} syllabification contrast.

An automated syllabification system was constructed and applied to medial clusters in the CELEX data. This system need only deal with simplex English words that contain the medial consonant clusters which are the focus of this study.  There is, for instance, no need to address the status of ``ambisyllabic'' consonants, i.e., singleton medial consonants preceded by a stressed lax vowel \citep[][219f.]{Rubach1996}, or to address morphological effects on syllabification. The technique used is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, 
\citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which parses word-medial clusters so that the onset is as large as possible. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), and onset maximization leaves the coda of the first syllable empty. In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} does not occur word-initially; here the maximal onset is [str], and the [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{propaga}[n.d]\emph{a}, \emph{whi}[s.p]\emph{er}, \emph{vi}[s.t]\emph{a}, or \emph{bi}[s.k]\emph{uit}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. \citet[][55]{Harris1994} notes that onset maximization makes the wrong predictions when the medial cluster is a valid onset: in \emph{whisper}, \emph{vista}, and \emph{biscuit}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is essentially that of \citet{Pulgram1970}: the first consonant of a medial cluster is assigned to the coda of a preceding coda if is immediately preceded by a stressed lax vowel, which do not occur syllable-finally in English.\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}

If the English affricates [tʃ, dʒ] are treated as sequences and not individual segments, this addition to the standard onset maximization technique would itself make the wrong prediction for medial affricates preceded by lax stressed vowels in words like \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}, incorrectly assigning the stop and fricative portions to separate syllables. For this reason, an additional constraint is assumed which prevents the two components of the affricates from being split by syllabification. This is motivated by the tendency of affricates to pattern with single segments in many languages. For instance, the only complex onsets in Classical Nahua are the affricates [ts, tʃ, tɬ] \citep[][9]{Launey2011}.

\subsection{Inventory consdierations and subsyllabic parsing}

To identify the inventory of syllable contact clusters, it is also necessary to parse syllables into onset, nucleus, and coda. In most cases this is trivial, but in a few cases, it is necessary to use distributional heuristics to determine the underlying inventory and subsyllabic affiliations.

\subsubsection{The velar nasal} \label{velarnasal}

There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ found before /k, g/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.). The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. \citet{Pierrehumbert1994} assumes the pure allophonic analysis in her study of syllable contact clusters, and it is assumed here. English nasal place allophony is formalized in Section \ref{cnpasection} below.

\subsubsection{[j] onglides}

The [j] onglide in words such as in words such as \emph{ass}[j]\emph{ume} is assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, J. \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is at least marginally contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}).

There is considerable evidence for this assumption. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters except in words like \emph{muse} or \emph{view}; the exceptionality of [juː] also suggests the glide is nuclear. There is also strong external support for this principle. The [juː] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005} and \emph{shm}-reduplication \citep{Nevins2003} to the exclusion of the rest of preceding tautosyllabic consonants. The same fusion of [juː] at the expense of the onset is also found in speech errors, e.g., [kjuː]\emph{mor} [h]\emph{omponent} for [hjuː]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.

\subsubsection{[w] onglides}

The selective properties of the back onglide [w] contrast sharply with those of the front onglide, and I assume that it is assigned to the onset. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. And onsets followed by [w] pattern with other complex onsets in undergoing epenthesis in the patholical speech of VBR, discussed immediately above.

\subsubsection{Post-vocalic \emph{r}}

Both CELEX and the dictionary used by \citet{Pierrehumbert1994} in her study transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost. In \emph{r}-full dialects, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters. Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}. Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which post-vocalic \emph{r} patterns differently than internal coda consonants. \citeauthor{Harris1994} reports that a variable process of /t/-\textsc{Glottalization} in many dialects of British English is blocked when /t/ is preceded by any consonant except post-vocalic \emph{r}.

\begin{example}[/t/-\textsc{Glottalization} in British English (after \citealp{Harris1994}:195, 258)]
\begin{tabular}{l l l@{} l}
a. & par[t]   &   & par[ʔ]   \\
   & car[t]on &   & car[ʔ]on \\
b. & fis[t]   & * & fis[ʔ]   \\
   & mis[t]er & * & mis[ʔ]er \\
\end{tabular}
\end{example}

\noindent Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.

\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
\begin{tabular}{l l l@{} l}
a. & be[lt]  &   & be[l] \\
   & me[nd]  &   & me[n] \\
b. & sh[ɚt]  & * & sh[ɚ] \\
   & c[ɚd]   & * & c[ɚ]  \\
\end{tabular}
\end{example}

@@

\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000}, \citealt{Dmitrieva2008a}, \citealt{Dmitrieva2008b} on English). The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented. Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.

@@

\subsection{Derived constraints}

@@

\begin{example}[Inflectional affix voice assimilation]
\begin{tabular}{l l l l l l}
a. & nap & nap[t] \\
   & nab & nab[d] \\
b. & lap & lap[s] \\
   & lab & lab[z] \\
\end{tabular}
\end{example}

@@

The failure of a small number of roots to conform to this generalization can be derived by marking such roots [$-$\textsc{Obstruent Voice Assimilation}]. Another possibility is that disagreeing obstruent clusters might be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}). There are precedents for using morphological boundaries solely to preserve phonological generalizations. In \emph{SPE} (p.~94), for instance, morphological boundaries are used to simplify stress assignment, and \citet[][546]{Rice2009d} analyses many words in the Athapaskan language Slave as compounds simply because they contain consonant clusters that rarely occur otherwise. Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}, and this ability seems to persist into adulthood \citep{McQueen1998}.

@@

\subsection{Static constraints}

@@

@@

\subsection{Discussion}

>>>>

%\chapter{Structural and accidental gaps\protect\symbolfootnote[1]{An abridged version of this chapter was presented at the 30th West Coast Conference on Formal Linguistics and will appear in the proceedings.}} \label{clusters}
\chapter{Structural and accidental gaps} \label{clusters}

This chapter

It has often been noted that these regularities, while perhaps statistical in nature,

\begin{quote}
These regularities, although not the products of alternations, are described and defined by elements of phonological theories: segments\ldots{}and distinctive features. \citep[13]{Berkley2000}
\end{quote}

@@

@@

One of the first study of this type was an analysis of co-occurrence restrictions on consonants in Javanese roots by \citet{Mester1988}, recently revisited by \citet{Graff2011}. 
Another Austronesian language studied in this fashion is Muna \citep{Coetzee2008a,Anttila2008}.
This technique has been applied to consonants in Semitic, especially Arabic 
(\citealt{McCarthy1988,McCarthy1994}, \citealt{Pierrehumbert1993}, \citealt{Frisch1996,Frisch2004}, \citealt{Coetzee2008a}) but also in Tigrinya \citep{Buckley1997}, Hebrew \citep{Berent2003}, and Amharic \citep{Colavin2010}, and to various co-occurrence restrictions in English (\citealt{Berkley1994b,Berkley1994a}, \citealt{Pierrehumbert1994}, \citealt{Coetzee2008b}), Russian \citep{Padgett1992}, Dutch \citep{Graff2011}, Navajo \citep{Martin2007,Martin2011} and Gitksan \citep{Brown2010}.

@@

\section{The static phonotactics hypothesis and the lexicon}

@@

The focus of this study are the constraints on syllable contact clusters (i.e., word-medial consonant clusters spanning syllable boundaries), also known as \emph{interludes}, in English monomorphs.

\citep{Haugen1956}
\footnote{Thanks to Stephen Anderson for bringing this study to my attention.}

\begin{quote}
\ldots{}in the absence of additional provisos, any concatenation of a well-formed coda and a well-formed onset is predicted to be possible medially in a word. \citep[][168]{Pierrehumbert1994}
\end{quote}

\noindent Before considering any ``additional provisos'', it is important to note that neither this study nor the study of \citet{Pierrehumbert1994} are not so much concerned so much with what clusters are ``possible'', but rather which are attested in the English lexicon.

One such proviso 
The neutralizing effect of phonological processes are one such proviso. Consider

\subsection{The focus on monomorphs}

This requires a bit of history.

A major result of American structuralism is that generalizations which hold of the sound sequences found in underlying forms may not hold across morph boundaries. In his history of phonology, \citet[][267]{Anderson1985} indirectly credits this insight to \citet{Bloomfield1930}, who proposes that a problem in the phonemic analysis of German can be resolved by allowing for allophony to make reference to morpheme boundaries. \citeauthor{Bloomfield1930} notes that [ç] and [x] can be cast as allophones, despite the existence of apparent minimal pairs like \emph{Kuchen} [kuːxən] `cake' $\sim$ \emph{Kuhchen} [kuːçən] `cow-let'. \citeauthor{Bloomfield1930}'s claim is that these words differ not just in their medial consonsant but also that the latter is a compound, marked with the diminutive \emph{-chen} (cf. \emph{Kuh} `cow'). Under the standard analysis that [x] is the allophone of /ç/ before back vowels, 
all that needs to be added is a restriction that the conditioning vowel belongs to the same word (or perhaps morph).

A more sophisticated approach to problems of this type, making no reference to \citeauthor{Bloomfield1930}, was published only two years later. \citet{Jakobson1932}, then a member of the Prague circle,

attributes a number of phonological properties of the Russian imperative to ``juncture''.

Nearly all Russian infinitives are marked with a final theme vowel and /-t\pal{}/; in reflexives, this is followed by a suffix realized as [s\pal{}ə], which triggers a general process of regressive voicing assimilation. In the reflexive infinitive, the palatalization of infinitive /-t\pal/ is lost, but in the reflexive imperative, palatalization of a stem-final consonant is preserved, even when the root-final consonant is /t\pal/, as in (\ref{rusreflex}c).\footnote{Thanks to Lev Blumenfeld for assistance with these transcriptions.}

\begin{example}[Russian reflexives] \label{rusreflex}
\begin{tabular}{l l l l}
   &  infinitive       & imperative \\
a. & [slav\pal{}itsə]  & [slaf\pal{}s\pal{}ə]  & `be glorious'   \\
   & [upram\pal{}itsə] & [upram\pal{}s\pal{}ə] & `be stubborn'   \\
b. & [kras\pal{}itsə]  & [kras\pal{}s\pal{}ə]  & `put on makeup' \\
   & [ʒar\pal{}itsə]   & [ʒar\pal{}{}s\pal{}ə] & `roast'         \\
c. & [zəbytsə]         & [zəbut\pal{}s\pal{}ə] & `forget'        \\
\end{tabular}
\end{example}

\noindent \citeauthor{Jakobson1932} proposes that the preservation of palatalization in /\ldots{}t\pal{}-s\ldots{}/ is a special property of the imperative. Both (\ref{rusreflex}c) forms contain a /t\pal-s/ juncture, though palatalization is only preserved in the imperative. \citeauthor{Jakobson1932} proposes that the phonology treats them differently: the imperative does not undergo loss of palatalization. A few years later, \citet{Trnka1936} makes the connection between \citeauthor{Jakobson1932}'s hypothesis and phonotactic generalizations.

This is apparent from a discussion of constraints on vowel sequences in Mixteco (a dialect cluster also known as Mixtec) given by \citet{Pike1947b}. It is apparent that \citeauthor{Pike1947b}'s constraints on vowel sequences are not intended to hold across morph boundaries, according to the phonological analysis of a Mixteco text published a few years earlier \citep{Pike1944}. A few examples are given below; the transcriptions are \citeauthor{Pike1944,Pike1947b}'s.

\begin{example}[Mixteco MSCs and complex words \citep{Pike1944,Pike1947b}]
\begin{tabular}{l l@{} l l l}
a. & * & {C}a{C}e & [ká\textsuperscript{n}dee]    & `kept inside'   \\
b. & * & {C}ə{C}e & [nìk\`əbəde]                  & `who entered'   \\
c. & * & {C}e{C}i & [teníke\textsuperscript{n}da] & `was walking    \\
d. & * & {C}i{C}e & [tenìkeet\`ə]                 & `and went away' \\
\end{tabular}
\end{example}

\noindent \citet[][166]{Pike1947b} affirms that the morpheme is ``marked'' by the violation of morpheme-internal sequence restrictions, an idea further developed by \citet{Harris1955} as a discovery procedure for morphs (see also \citet{Hafer1974}).

and lead children \citep{Mattys2001b,Jusczyk2002,Davidson2004} and adults \citep{McQueen1998,Hay2004a} to posit morph boundaries (see also \S\ref{harmonyandsegmentation}).

% 
%

@@

a categorization of morphemic complexity 
and broad transcriptions

@@

\subsubsection{Identifying monomorphs}

% Pierrehumbert informal judgements * \citet{Pierrehumbert1994} and \citet{Hammond1999a} technique is not replicable, so we turn to CELEX
% not replicable, either the experience nor the data itself
% Some of Pierrehumbert's and some of Hammond's are ridiculous
% Despite this, we preserve some Pierrehumbertian generalizations

Beyond this, there is a potential for bias or circularity of argumentation. As mentioned above (\S\ref{pike}), a sequence which is rare in monomorphs is sometimes said to mark a juncture even in the absense of other evidence for segmentation. But it is certainly incorrect to use this principle to argue for a juncture if the argument is only being made to preserve this very phonotactic generalization, as \citet[186]{Pierrehumbert1994} does with the word \emph{antler}.

% Compounds, regulars, and more productive derivatives are pretty clear 
% Latin is less clear: \citet{Bolinger1961} on stress, and \emph{SPE} and \citet[11--13]{Aronoff1976} on forms of prefixes and bound stems
%There has been a good deal of attention paid to Latinate prefixes
%\citet{Bolinger1961} suggests this for English stress.
%\citet{SPE}

\begin{example}[Bound stem-specific allomorphy]
\begin{tabular}{l l l}
a. & \emph{adhere}   & \emph{adhesion}   \\
   & \emph{cohere}   & \emph{cohesion}   \\
b. & \emph{conceive} & \emph{conception} \\
   & \emph{perceive} & \emph{perception} \\
\end{tabular}
\end{example}

Note that this is common behavior in other languages in which the independent nature of bound stems is less controversial, such as Russian \citep{Pesetsky1977}.

% Harley2009 \citet{Gropen1989} \citet[56]{Coppock2008} observes that there is every reason to believe that the ditransitive is productive in the sense that it can be applied to new words; consider recent conversion \emph{text} (e.g., \emph{text me your number}).

\begin{example}[Latinate verbs and small clauses]
\begin{tabular}{l l@{} l l l@{}}
a. & ~ & {show it off}           & * & {exhibit it off}           \\
b. & ~ & {drink himself stupid}  & * & {imbibe himself stupid}    \\
c. & ~ & {show him the painting} & * & {exhibit him the painting} \\
\end{tabular}
\end{example}

% While this is precisely the abstractness that  ???, it is reasonably well supported by psycholinguistic results in lexical decision research.

% description of taft model \citet{Taft1975} propose an influential model of morphological processing. The first step in the \citeauthor{Taft1975} model is to strip affixes from possible roots. Root and affix are then looked up in lexical memory. At the final stage, they are recombined. 
% It has been suggested since then that full decomposition is only one of two ways to recognize complex words, the other being a whole word storage route running in cascade \citep{Caramazza1988} or in parallel \citep{Baayen1997b}. Whatever evidence is amassed for this (see \citealp{Lignos2012} for a critical review) but few if any lexical decision researchers would deny the possibility of \citep{Butterworth1983}.

% first experiment \citet{Manelis1977} who suggest poor stimulus controls \citeauthor{Taft1975}

% that *\emph{devive} in the Latinate prefix \emph{de-} applies to a bound stem \emph{vive} (e.g., \emph{revive}) is slower to be recognized than *\emph{debel}, since \emph{bel} is not recognized as abound stem in words like \emph{rebel}.

% pseudoprefixed 
% \citet{Taft1981} 
% \citet{Rubin1979}
%\emph{regatta}
%\emph{retrieve}

%\citet{Taft2004b}
%\emph{virus} is primed by \emph{viral}

\subsubsection{Lexical statistics}

The English portion of CELEX contains 6,876 simplex words. Under the syllabification procedure described immediately below, this yields contain 21 medial coda types and 40 unique medial onsets. Of the 840 ($= 21 \times 40$) possible syllable contact clusters that could be produced by free combinations of the attested medials codas and onsets, 158 are attested, an 18.8\% saturation rate.

Cluster frequencies are reproduced in Appendix \ref{appendixB}.

\subsubsection{Syllabification}

CELEX provides transcriptions of individual wordforms, include syllabification. However, the authors do not describe the syllabification procedure, and the syllable boundaries appear to be placed in an unsystematic fashion. For instance, \emph{chemistry} is syllabified as [ˈkɛ.mɪ.strɪ] but \emph{ministry} as [ˈmɪ.nɪs.trɪ].\footnote{Note that word-final \emph{y} is generally lax [ɪ] in Received Pronunciation \citep[][II.294]{AOE}.} This putative [ɪ.strɪ $\sim$ ɪs.trɪ] contrast does not conform to native speaker intuition, and it is typologically dubious: there is no convincing evidence for contrastive syllabification in any language. Apparent counterexamples appear to be interactions of syllabification with vowel or consonant length \citep[e.g.,][]{Elfner2006}, word stress, or morphological structure, none of which might predict a different syllabification for \emph{chemistry} and \emph{ministry}. Many other instances of contrasts in syllabification can be found in the CELEX data.

For this reason, an automated syllabification procedure was developed to separate medial clusters in the CELEX  monomorphs into coda and onset. Since this procedure is applied only to medial clusters, it is possible to remain agnostic on several contentious issues concerning syllabification of English; for instance, there is no need to address the status of so-called ``ambisyllabic'' consonants, since this does not occur with medial clusters, or morphological effects on syllabification, since only monomorphs are used.

The automated syllabifier is a variation on the theme of onset maximization (e.g., \citealt{Kurylowicz1948}, \citealt[75]{Pulgram1970}, \citealt[42f.]{Kahn1976}, \citealt[][358f.]{Selkirk1982b}) which favors parses of word-medial clusters in which as much of the cluster as possible is assigned to the onset. Just what is a ``possible'' medial onset is defined by the onsets that occur in initial position. While this equation may not be appropriate for all languages \citep[306]{Fischer-Jorgensen1952}, it is for English. Medial clusters in words like \emph{neu}[.tr]\emph{on} or \emph{bi}[.str]\emph{o} are found in word-initial position (e.g., [tr]\emph{ansit}, [str]\emph{ike}), so onset maximization assigns the entire medial cluster to the onset, leaving the medial coda empty. In contrast, the [nstr] cluster in \emph{mi}[n.str]\emph{el} is not found word-initially; the maximal onset here is [str], and the residual [n] is assigned to the coda.

There is one case where unchecked maximization of medial onsets produces incorrect syllabifications. When a medial consonant cluster is preceded by a stressed lax vowel, as in words like \emph{wh}[ɪs.p]\emph{er}, \emph{v}[ɛs.t]\emph{ige}, or \emph{m}[ʌs.k]\emph{et}, the first consonant of the cluster checks the lax vowel \citep[e.g.,][3]{Hammond1997}. \citet[][55]{Harris1994} note that onset maximization produces an incorrect syllabification when the medial cluster is a valid onset: in \emph{whisper}, \emph{vestige}, and \emph{musket}, it incorrectly assigns [sp, st, sk] to the medial onset, leaving the coda empty. The approach adopted here is is to assign the first consonant of a medial cluster to the coda of a preceding syllable if it follows a stressed lax vowel \citep[cf.][48]{Pulgram1970}.%\footnote{\citet[][589f.]{Lowenstamm1981} observes a strikingly similar issue with the onset maximization principle in French, and proposes a similar solution.}

\paragraph{The status of affricates} At least as far back as the 1939 first edition of the \emph{Grundz{ü}ge der Phonologie} \citep{Trubetzkoy1958}, affricates have been analyzed as single plosive segments distinguished from stops by the presence of stridency \citep[24]{Jakobson1961} or a delayed release (\emph{SPE}:321f). If however, the English affricates [tʃ, dʒ] are treated as stop-fricative sequences (e.g., \citealt{Hualde1988}, \citealt{Lombardi1990}), this procedure will incorrectly assign the two components of the medial affricate to separate syllables before stressed lax vowels (e.g., \emph{ra}[.tʃ]\emph{et} or \emph{a}[.dʒ]\emph{ile}) unless there is an ad-hoc constraint that treats these sequences as indivisible \citep[e.g.,][]{Wells1990}. Here, the traditional treatment of affricates as single segments is assumed. This assumption can be motivated by the tendency of affriates to pattern with individual segments in phonotactic generalizations. For instance, Classical Nahua allows the affricate series [ts, tʃ, tɬ] in onsets, but otherwise does not permit other onset clusters \citep[9]{Launey2011}, and \citet[86]{Butskhrikidze2002} notes a similar generalization in the consonant clusters of Georgian.

\subsubsection{Subsyllabic parsing}

It is not trivial in all cases to separate interludes from the nuclei on either side. Immediately below, distributional heuristics are used to develop a subsyllabic parsing procedure for English.

\paragraph{The velar nasal} There is a long-standing debate regarding whether English [ŋ] is a phoneme in its own right \citep[e.g.,][]{Sapir1925}, an analysis which is a consequence of principles like the Alternation Condition \citep{Kiparsky1968} or Lexicon Optimization \citep[][53]{OT}, or simply the allophone of /n/ when followed by underlying /k, ɡ/ (e.g., \emph{SPE}:85, \citealt{Borowsky1986}:65f.), sometimes followed by the deletion of /ɡ/. The strongest piece of evidence for the latter position, in which the velar nasal is a pure allophone, is the general absence of onset [ŋ], a position where it can never be followed by a dorsal consonant needed to derive the velar allophone. Following \citet{Pierrehumbert1994}, the pure allophonic analysis is adopted here, and English nasal place allophony is formalized in Section \ref{npasection} below.

\paragraph{[j] onglides} The [j] onglide in words such as \emph{val}[j]\emph{ue} are assumed to be present in underlying representations (e.g., \citealt[][278]{Borowsky1986}, \citealt{Anderson1988b}, pace \emph{SPE}:196, \citealt[][89]{Halle1985a}, \citealt[][217]{McMahon1990}), since the presence or absence of the glide is contrastive (e.g., \emph{coo}/\emph{coup} \alt{} \emph{queue}, \emph{booty} \alt{} \emph{beauty}). The front onglide is further assumed to be assigned to the nucleus, except when the onset would otherwise be null (e.g., \emph{jun}[j]\emph{or}). There is considerable evidence for this assumption. When [j] is a simplex onset, it may be followed by any vowel \citep[][276]{Borowsky1986}, but when [j] is immediately preceded by an onset consonant (e.g., [bj]\emph{ugle}), the following vowel is always [uː]. 
%In fact, speakers judge words in which the front onglide is preceded by an onset consonant and followed by a vowel other than [uː] to be anomalous \citep{Moreland2009}. 
This defective distribution of vowels following [j] suggests that the onglide in this context is the first component of a phonological diphthong (\citealp[][232]{Hayes1980}, \citealp[][61f.]{Harris1994}, \citealp{Davis1995}). \citet[][42]{Clements1983} note that /m, v/ do not appear in onset clusters, though they may be followed by [ju] in words like [mj]\emph{use} or [vj]\emph{iew}, also suggesting the nuclear affiliation of the glide. There is also strong external support for this principle. The [ju] in words like \emph{spew} may pattern together in Pig Latin \citep{Davis1995,Idsardi2005}, \emph{shm}-reduplication \citep{Nevins2003}, and speech errors, such as [kjuː]\emph{mor} [h]\emph{omponent} for intended [hju]\emph{mor component} \citep[][130]{Shattuck-Hufnagel1986}.

\paragraph{[w] onglides} The selective properties of the back onglide [w] contrast sharply with those of the front onglide. Whereas the front onglide shows only limited selectivity for preceding tautosyllabic consonants \citep{Davis1995,Kaye1996}, the back onglide [w] is rarely preceded by tautosyllabic consonants other than [k] (e.g., \emph{tran}[kw]\emph{il}). Unlike the front glide, syllable-initial [kw] may be followed by nearly any vowel \citep[][161]{Davis1995}. Unlike [juː], onglide [w] followed by a vowel does not pattern together in Pig Latin \citep[][166]{Davis1995}. Together, this indicates taht the back onglide is assigned to the onset.

\paragraph{Post-vocalic \emph{r}} Both CELEX and the dictionary used by \citet{Pierrehumbert1994} transcribe Received Pronunciation, in which word-medial post-vocalic \emph{r} has been lost. Even in \emph{r}-full dialects, though, there is reason to believe that post-vocalic \emph{r} is in fact nuclear, and therefore its presence or absence is irrelevant to the constraints on syllable contact clusters. Many vowel contrasts are suspended before \emph{r} (e.g., \citealt[269f.]{Fudge1969}, \citealt[][255]{Harris1994}); for example, \emph{fern}, \emph{fir}, and \emph{fur} lack the contrasts found in \emph{pet}, \emph{pit}, and \emph{putt}. Further evidence for the nuclear status of post-vocalic \emph{r} comes from variable phonological processes in which it patterns differently than internal coda consonants. \citet[258]{Harris1994} reports that post-vocalic \emph{r} is the only consonant which does not block a variable process of /t/-\textsc{Glottalization} found in many British dialects.

\begin{example}[/t/-\textsc{Glottalization} in British English]
\begin{tabular}{l l l l@{} l}
a. & \emph{des}[ɚt]         & \alt{} &   & \emph{des}[ɚʔ]         \\
   & \emph{c}[ɚt]\emph{ain} & \alt{} &   & \emph{c}[ɚʔ]\emph{ain} \\
b. & \emph{fi}[st]          & \alt{} & * & \emph{fi}[sʔ]          \\
   & \emph{mi}[st]\emph{er} & \alt{} & * & \emph{mi}[sʔ]\emph{er} \\
\end{tabular}
\end{example}

Similarly, while /t, d/ delete in word-final position when immediately preceded by a consonant, including sonorants /n, l/, as in (\ref{td}a), deletion of /t, d/ after post-vocalic \emph{r} in American English is ``rare or nonexistent'' \citep[][8]{Guy1980}.

\begin{example}[/t, d/-\textsc{Deletion} in American English] \label{td}
\begin{tabular}{l l l l@{} l}
a. & \emph{be}[lt] & \alt{} &   & \emph{be}[l] \\
   & \emph{me}[nd] & \alt{} &   & \emph{me}[n] \\
b. & \emph{sh}[ɚt] & \alt{} & * & \emph{sh}[ɚ] \\
   & \emph{c}[ɚd]  & \alt{} & * & \emph{c}[ɚ]  \\
\end{tabular}
\end{example}

@@

\citeauthor{Mester1988}'s technique has been adopted by many subsequent studies (e.g., \citealt{Padgett1992,Padgett1995} on Russian; \citealt{Pierrehumbert1993}, \citealt{Frisch1996}, and \citealt{Frisch2004} on Arabic; \citealt{Berkley1994b,Berkley1994a,Berkley2000} on English).
The null hypothesis $H_0$ is that some sequence occurs at no different a rate than the researcher expects, and the alternative hypothesis $H_1$ is that it is underrepresented. Given an observed frequency $O$ and also a frequency $E$ expected under the null hypothesis, the test is as follows.

@@

@@

\begin{example}[Inflectional affix voice assimilation]
\begin{tabular}{l l l l l l}
a. & \emph{nap} & \emph{nap}[t] \\
   & \emph{nab} & \emph{nab}[d] \\
b. & \emph{lap} & \emph{lap}[s] \\
   & \emph{lab} & \emph{lab}[z] \\
\end{tabular}
\end{example}

@@

The failure of a small number of roots to conform to this generalization can be derived by marking such roots [$-$\textsc{Obstruent Voice Assimilation}]. Another possibility is that disagreeing obstruent clusters might be analyzed as semantically opaque compounds by native speakers (e.g., \emph{ja}[k\#d]\emph{aw}). There are precedents for using morphological boundaries solely to preserve phonological generalizations. In \emph{SPE} (p.~94), for instance, morphological boundaries are used to simplify stress assignment, and \citet[][546]{Rice2009d} analyses many words in the Athapaskan language Slave as compounds simply because they contain consonant clusters that rarely occur otherwise. Recent experiments find that 9-month-old infants are sensitive to the contrast between [vt], a disagreeing cluster which does not occur word-medially, and [ft], found in medial position \citep{Mattys2001b}. And this ability seems to persist into adulthood (\citealt{Brown1956}, \citealt{Norris1997}, \citealt{McQueen1998}).

@@

\subsubsection{Local summary}

All three \emph{SPE} rules affecting syllable consonant clusters are reliable predictors of attested and unattested medial clusters.

\subsubsection{Static constraints}

@@

\subsubsection{Local summary: static constraints}

The inventory of attested clusters shows no statistically reliable traces of the three static constraints proposed by \citet{Pierrehumbert1994}.

@@

\section{Conclusions}
